Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-19 ba_resubmit/kundinger_resubmit2.tex
small comment.

		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations. Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.  Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B and B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.  For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records. For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.

		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison vectors} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
		the 
\textcolor{teal}{memory-intensive}
		\textcolor{teal}{comparison vectors} $\gamma^{ab}$ 
		before continuing with the next chunk of data. 
		In detail, we calculate
\begin{subequations}
	\begin{align}
		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
	\end{align}
\end{subequations}
These can computed \textcolor{teal}{sequentially} or in parallel. Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
\begin{subequations}
	\begin{align}
		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
\end{subequations}

@@

As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$. However, the memory requirements of each are still quadratic in nature. For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome. We next introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches. This allows all-to-all comparisons for substantially larger linkage tasks.

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $SEI(r_{p_j}^{ab})$. We collect these memory reduced lists to form $SEI(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $SEI(\mathcal{R})$.

Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively. Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels. As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $SEI(\tilde{\gamma}) = \{\mathcal{P}, SEI(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section \ref{SEI-sensitivity}.

>>>>

	\textcolor{red}{Suggestion: Chunk is award. Perhaps partition/block/batch/segment. I might like segment since partition/block have meanings in record linkage already. }

		For linkage tasks with large amounts of records, we can partition $A$ and $B$ into $t_A$ and $t_B$ smaller disjoint chunks for more manageable computations. Let $\{A^1, \dots, A^{t_A}\}$ be a partition of $A$ such that $\cup_{a=1}^{t_A} A^a = A$ and $A^a \cap A^{a'}=\emptyset$ for all $a \neq a'$.

		Likewise, let $\{B^1, \dots, B^{t_b}\}$ be a partition of $B$ such that $\cup_{b=1}^{t_B} B^b = B$ and $B^b \cap B^{b'}=\emptyset$ for all $b \neq b'$.  For example, if $n_A = n_B = 50000$ and $t_A = t_B = 100$, each $A^a$ and $B^b$ would contain disjoint sets of 500 records. For each $A^a$, we conduct all-to-all comparisons with each $B^b$ to construct the comparison matrix $\gamma^{ab}$.

		\textcolor{teal}{Then, we} conduct hashing, obtain the compressed \textcolor{teal}{comparison matrix} $\tilde{\gamma}^{ab}$, and \textcolor{teal}{remove}
		the 
\textcolor{teal}{memory-intensive}
		\textcolor{teal}{comparison matrix} $\gamma^{ab}$ 
		before continuing with the next chunk of data. 
		In detail, we calculate
\begin{subequations}
	\begin{align}
		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, j \in B^b\}, \label{r_{p_j}^{ab}} \\
		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
	\end{align}
\end{subequations}
These can computed \textcolor{teal}{sequentially} or in parallel. Summary statistics from each pairwise chunk comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
\begin{subequations}
	\begin{align}
		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
\end{subequations}

@@

As discussed in Section \ref{data-representation-hashing-and-storage}, storing the indices, patterns, and counts in $\tilde{\gamma}$ uses less memory than storing the full comparison matrix $\gamma$. However, the memory requirements of each are still quadratic in nature. For very large linkage tasks, recording the indices for all record pairs in $\mathcal{R}$ can become computationally burdensome. We introduce storage efficient indexing (SEI), which, when used with the methods in Section \ref{scaling}, allows us to compute $\mathcal{N}$ for all $n_A n_B$ record pairs while greatly reducing the memory costs of $\mathcal{R}$ associated with unlikely matches. This allows all-to-all comparisons for substantially larger linkage tasks.

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, we know that records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form $\text{SEI}(r_{p_j}^{ab})$. We collect these memory reduced lists to form $\text{SEI}(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form $\text{SEI}(\mathcal{R})$.

Let $n_{A^a}$ and $n_{B^b}$ be the number of records in chunks $A^a$ and $B^b$ respectively. Instead of storing $n_A n_B$ record labels, with SEI we store at most $\sum_{a = 1}^{t_A}\sum_{b = b}^{t_A} n_{A^a}  n_{B^b}  P  S$ labels. As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced $\text{SEI}(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
